---
title: "Data structure and Characteristics"
author: "Matteo Delucchi"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 3
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{Data structure and Characteristics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(dplyr)
library(ggplot2)
library(kableExtra)

# Save plots as files? 
# Warning: Bad layout with knitr. Works well if chunks are run without rendering.
SAVEPLOTS <- FALSE
PLOTPATH <- Sys.getenv("PLOTPATH")
```

# Load Preprocessed Data

```{r}
str(adb.raw)
```

# Characteristics of whole, harmonized data cohort

no. of patients
```{r}
nrow(adb.raw)
```
how many had a form of stroke as reason for diagnosis?
```{r}
table(adb.raw$Basis.of.recruitment_CaseControl, useNA = "always")
```

median age at diagnosis?
```{r}
summary(adb.raw$age.at.time.of.diagnosis)
```

median age at rupture?
```{r}
summary(adb.raw$age.at.time.of.diagnosis[which(adb.raw$Ruptured_IA=="Yes")])
```

Compare to Morel et al. 2021: 
Median age at aneurysm rupture was 52 years (range 10-92), 
25% of patients were <44 and 75% were <61 years of age.


# Variable Description

```{r}
abndata <- exp11_dat$abndata %>%
  # Clean up node labels
  rename("Sex" = "Gender",
         "Age at Diagnosis" = "AgeDiag.group",
         # "Age at Diagnosis" = "AgeDiag",
         "Pos. Fam. History" = "Positive.famillial.history",
         "Smoking Status" = "Smoking_Current_Former_No",
         # "Smoking Status=Former" = "Smoking_Current_Former_No=Former",
         # "Smoking Status=No" = "Smoking_Current_Former_No=No",
         "IA Location" = "location.grouped",
         # "IA Location=Low risk" = "location.grouped=Low",
         # "IA Location=Medium risk" = "location.grouped=Medium",
         # "IA Location=High risk" = "location.grouped=High",
         "Multiple IAs" = "Multiple.IAs",
         # "IA Size" = "IAsize_log", 
         "IA Size" = "IAsize.groups.merged", 
         "Ruptured IA" = "Ruptured_IA")
str(abndata)


```

```{r}
text_tbl <- data.frame(
  variables = c(
    "Family History",
    "Sex",
    "Age",
    "",
    "Smoking",
    "Hypertension",
    "Location",
    "Multiplicity",
    "Size",
    "",
    "Ruptured"),
  modvars = c(
    "discrete",
    "discrete",
    "continuous",
    "discrete",
    "discrete",
    "discrete",
    "discrete",
    "discrete",
    "continuous",
    "discrete",
    "discrete"
  ),
  descr = c(
      "Positive familial history ('Yes', 'No') of intracranial aneurysm if one or more first degree relative(s) were diagnosed with IA.",
      "Biological, self-reported sex ('female', 'male').",
      "Age in years at the time of IA diagnosis. Equal to age at time of rupture for ruptured IA.",
      "Age discretized in groups of A=[0,40), B=[40,45), C=[45,50), D=[50,55), E=[55,60), F=[60,65), G=[65,93).",
      "Patient's smoking behaviour classified as current smoker if smoked (> 300 cigarettes) and continues current smoking. Former smoker, if smoked (>300 cigarettes) and stopped at least 6 months pre-diagnosis. Never smoked if < 300 cigarettes smoked ever.",
      "No hypertension if patient is unaware of increased blood pressure (>120/80mmHg). Hypertension of AnyType if patient is aware of increased blood pressure. Including untreated and treated BP >140/90mmHg, with the later including both, cases of uncontrolled and controlled BP where the BP is above or in normal range respectively.",
      "IA Location classified by rupture risk in groups of high risk (Acom, Pcom, V-B, A2, PC), medium risk (MCA, ICA, Basilar, Other, A1 segment ant) and low risk (CavICA, OphtICA).",
      "Yes if >1 IA was detected.",
      "Natural logarithm of IA max diameter [mm] reported at time of diagnosis.",
      "IA max diameter [mm] grouped in in A=< 7mm, B=7-12mm, C=13-25mm, D=>25mm.",
      "IA rupture status at time of diagnosis.")
)
```


```{r}
text_tbl %>%
  rename("Factor" = "variables",
         "Type" = "modvars",
         "Description" = "descr") %>%
  
  kbl(format = "latex", booktabs = T) %>%
  # kable_classic(full_width = F) %>%
  kable_styling(latex_options=c("striped", "hold_position"),
                stripe_index = c(1,3,4,6,8,11)) %>%
  column_spec(3, width = "10cm") %>%
  # kableExtra::landscape() %>%
  cat(., file = paste0(PLOTPATH, "/vardesc_table.tex"))



  ableExtra::pack_rows("Age", start_row = 4, end_row = 5, bold = F) %>%
  kableExtra::pack_rows("Size", start_row = 10, end_row = 11, bold = F)
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "30em", background = "yellow")
```

