---
title: "Data structure and Characteristics"
author: "Matteo Delucchi"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 3
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{Data structure and Characteristics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(dplyr)
library(ggplot2)
library(kableExtra)

# Save plots as files? 
# Warning: Bad layout with knitr. Works well if chunks are run without rendering.
SAVEPLOTS <- FALSE
PLOTPATH <- Sys.getenv("PLOTPATH")
```

# Load Preprocessed Data

```{r}
str(adb.raw)
```

# Characteristics of whole, harmonized data cohort

no. of patients
```{r}
nrow(adb.raw)
```
how many had a form of stroke as reason for diagnosis?
```{r}
table(adb.raw$Basis.of.recruitment_CaseControl, useNA = "always")
```

median age at diagnosis?
```{r}
summary(adb.raw$age.at.time.of.diagnosis)
```

median age at rupture?
```{r}
summary(adb.raw$age.at.time.of.diagnosis[which(adb.raw$Ruptured_IA=="Yes")])
```

Compare to Morel et al. 2021: 
Median age at aneurysm rupture was 52 years (range 10-92), 
25% of patients were <44 and 75% were <61 years of age.


# Variable Description

```{r}
abndata <- exp11_dat$abndata %>%
  # Clean up node labels
  rename("Sex" = "Gender",
         "Age at Diagnosis" = "AgeDiag.group",
         # "Age at Diagnosis" = "AgeDiag",
         "Pos. Fam. History" = "Positive.famillial.history",
         "Smoking Status" = "Smoking_Current_Former_No",
         # "Smoking Status=Former" = "Smoking_Current_Former_No=Former",
         # "Smoking Status=No" = "Smoking_Current_Former_No=No",
         "IA Location" = "location.grouped",
         # "IA Location=Low risk" = "location.grouped=Low",
         # "IA Location=Medium risk" = "location.grouped=Medium",
         # "IA Location=High risk" = "location.grouped=High",
         "Multiple IAs" = "Multiple.IAs",
         # "IA Size" = "IAsize_log", 
         "IA Size" = "IAsize.groups.merged", 
         "Ruptured IA" = "Ruptured_IA")
str(abndata)


```

```{r}
text_tbl <- data.frame(
  variables = c(
    "Sex",
    "Age",
    "",
    "Family History",
    "Hypertension",
    "Smoking",
    "Multiplicity",
    "Location",
    "Size",
    "",
    "Ruptured"),
  modvars = c(
    "discrete",
    "discrete",
    "continuous",
    "discrete",
    "discrete",
    "discrete",
    "discrete",
    "discrete",
    "continuous",
    "discrete",
    "discrete"
  ),
  descr = c(
      "Biological, self-reported sex as \textit{female} or \textit{male}.",
      "Age in years at the time of IA diagnosis. Equal to age at time of rupture for ruptured IA.",
      "Age discretized in groups of A=20-39, B=40-44, C=45-49, D=50-54, E=55-59, F=60-64, G=65-93.",
      "Positive familial history of intracranial aneurysm as \textit{yes} if one or more first degree relative(s) were diagnosed with IA and \textit{no} otherwise.",
      "\textit{No} hypertension awareness if patient is unaware of increased blood pressure (>120/80mmHg). \textit{Yes} if patient is aware of increased blood pressure. Including untreated and treated BP >140/90mmHg, with the later including both, cases of uncontrolled and controlled BP where the BP is above or in normal range respectively.",
      "Patient's smoking behaviour classified as \textit{current} smoker if smoked (> 300 cigarettes) and continues current smoking. \textit{Former} smoker, if smoked (>300 cigarettes) and stopped at least 6 months pre-diagnosis. \textit{Never} smoked if < 300 cigarettes smoked ever.",
      "\textit{Yes} if >1 IA was detected.",
      "IA Location classified by rupture risk in groups of \textit{high} risk (Acom, Pcom, V-B, A2, PC), \textit{medium} risk (MCA, ICA, Basilar, Other, A1 segment ant) and \textit{low} risk (CavICA, OphtICA).",
      "Natural logarithm of IA max diameter [mm] reported at time of diagnosis.",
      "IA max diameter [mm] grouped in in \textit{A}≤ 7mm, \textit{B}=7-12mm, \textit{C}=13-25mm, \textit{D}≥25mm.",
      "\textit{Yes} if IA was ruptured at time of diagnosis and \textit{no} otherwise.")
)
```


```{r}
text_tbl %>%
  rename("Factor" = "variables",
         "Type" = "modvars",
         "Description" = "descr") %>%
  
  kbl(format = "latex", booktabs = T) %>%
  # kable_classic(full_width = F) %>%
  kable_styling(latex_options=c("striped", "hold_position"),
                stripe_index = c(1,4,6,8,11)) %>%
  column_spec(3, width = "10cm") %>%
  # kableExtra::landscape() %>%
  cat(., file = paste0(PLOTPATH, "/vardesc_table.tex"))



  ableExtra::pack_rows("Age", start_row = 4, end_row = 5, bold = F) %>%
  kableExtra::pack_rows("Size", start_row = 10, end_row = 11, bold = F)
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "30em", background = "yellow")
```

