---
title: "Model Inference of discrete BNs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Inference of discrete BNs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(dplyr)
library(bnlearn)


# Save plots as files? 
SAVEPLOTS <- F
PLOTPATH <- paste0(Sys.getenv("PLOTPATH"), "/fourierplots")
PLOTFORMAT <- "pdf" # "pdf", "svg" or "png"
PLOTWIDTH = 16
PLOTHEIGHT = 9
```

# Load discrete additive mixed-effects BN

```{r message=FALSE, warning=FALSE}
load(paste0(PLOTPATH, "/discrete_BN_SL_additive_bic_mixedeffects.RData"))
```


```{r}
data <- tobesaved$madbn_data %>%
  mutate(positive_famillial_history = factor(positive_famillial_history))
bl <- tobesaved$madbn_bl
avg <- tobesaved$madbn_avg
bb <- tobesaved$madbn_boot
```

```{r}
strength.plot(avg,
              bb,
              main = "discrete additive mixed-effects BN",
              # sub = "SL: tabu\nth.=0.3\nScore=BIC",
              shape = "rectangle")
```



# Inference by querying


```{r}
madbn_fit <- bn.fit(data = data, x = avg)
```

```{r}
# remove study_source
avg_mod <- bnlearn::remove.node(avg, "study_source")
bb_mod <-  bb %>%
  filter(from != "study_source") %>%
    filter(to != "study_source")
attr(bb_mod, "nodes") <- attr(bb_mod, "nodes")[-10]
data_mod <- data %>%
  select(-c("study_source"))

strength.plot(avg_mod,
              bb_mod,
              main = "discrete additive mixed-effects BN",
              sub = "Fitting w/o study_source",
              shape = "rectangle")

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_admeBN_fitting_wo_studysource"  
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } 
} 

madbn_fit_mod <- bn.fit(data = data_mod, x = avg_mod)
```


```{r export as HUGIN flat network format}
bnlearn::write.net(file = paste0(PLOTPATH, "/adme_fitted.net"), fitted = madbn_fit)
bnlearn::write.net(file = paste0(PLOTPATH, "/adme_fitted_wo_studysource.net"), fitted = madbn_fit_mod)
```



```{r}
plot(madbn_fit_mod$hpt_aware$prob)
```

Males tend to have an IA more often at a high risk location but
Females tend to develop an IA in high risk location more often if they suffer HBP than males who suffer HBP.


```{r}
plot(madbn_fit_mod$IAsize_diag_grouped_merged$prob)
```

Being aware of pos. fam. hist. reduces the porbability of developing larger IAs.

```{r}
plot(madbn_fit_mod$IAlocation_group$prob)
```


```{r}
plot(madbn_fit_mod$IAruptured$prob)
```

Larger IAs tend to rupture more frequently.

```{r }
par(mfrow = c(2,2))
sim = cpdist(madbn_fit, nodes = c("hpt_aware", "age_diag_group"), n = 10^4, evidence = (IAruptured == "yes"))
plot(sim, col = "grey",
     main = "IAruptured == yes")

sim = cpdist(madbn_fit, nodes = c("hpt_aware", "age_diag_group"), n = 10^4, evidence = (IAruptured == "no"))
plot(sim, col = "grey",
     main = "IAruptured == no")

sim = cpdist(madbn_fit, nodes = c("age_diag_group", "hpt_aware"), n = 10^4, evidence = (IAruptured == "yes"))
plot(sim, col = "grey",
     main = "IAruptured == yes")

sim = cpdist(madbn_fit, nodes = c("age_diag_group", "hpt_aware"), n = 10^4, evidence = (IAruptured == "no"))
plot(sim, col = "grey",
     main = "IAruptured == no")

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-age-hpt"  
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```


```{r }
par(mfrow = c(1,2))
sim = cpdist(madbn_fit, nodes = c("age_diag_group"), n = 10^5, evidence = (IAruptured == "yes"))
plot(prop.table(table(sim)), col = "grey",
     main = "IAruptured == yes")

sim = cpdist(madbn_fit, nodes = c("age_diag_group"), n = 10^5, evidence = (IAruptured == "no"))
plot(prop.table(table(sim)), col = "grey",
     main = "IAruptured == no")

# SAVEPLOTS <- F
if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-age"  
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
} 
```


```{r}
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "IAsize_diag_grouped_merged"), n = 10^5, evidence = (IAruptured == "yes"))
plot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "yes"')

sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "IAsize_diag_grouped_merged"), n = 10^5, evidence = (IAruptured == "no"))
plot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "no"')

sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "IAsize_diag_grouped_merged"), n = 10^5, evidence = (IAruptured == "yes"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "yes"')

sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "IAsize_diag_grouped_merged"), n = 10^5, evidence = (IAruptured == "no"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "no"')

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-smoking-size"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```


```{r}
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "age_diag_group"), n = 10^5, evidence = (IAruptured == "yes"))
plot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "yes"')
round(prop.table(table(sim)), 2)

sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "age_diag_group"), n = 10^5, evidence = (IAruptured == "no"))
plot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "no"')
round(prop.table(table(sim)), 2)

sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "age_diag_group"), n = 10^5, evidence = (IAruptured == "yes"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "yes"')
round(prop.table(table(sim)), 2)

sim = cpdist(madbn_fit, nodes = c("smoking_current_notcurrent", "age_diag_group"), n = 10^5, evidence = (IAruptured == "no"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'IAruptured == "no"')
round(prop.table(table(sim)), 2)

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-smoking-age"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```


```{r}
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "age_diag_group"), n = 10^5, evidence = (smoking_current_notcurrent == "current" & IAruptured == "yes"))
plot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "current" & IAruptured == "yes"')

sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "age_diag_group"), n = 10^5, evidence = (smoking_current_notcurrent == "current" & IAruptured == "no"))
plot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "current" & IAruptured == "no"')

sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "age_diag_group"), n = 10^5, evidence = (smoking_current_notcurrent == "current" & IAruptured == "yes"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "current" & IAruptured == "yes"')

sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "age_diag_group"), n = 10^5, evidence = (smoking_current_notcurrent == "current" & IAruptured == "no"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "current" & IAruptured == "no"')

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_smoking-age-size"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```

```{r}
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("gender", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
plot(prop.table(table(sim)),
     main = 'IAruptured == "yes"')
round(prop.table(table(sim)), 2)
           
sim = cpdist(madbn_fit, nodes = c("gender", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
plot(prop.table(table(sim)),
     main = 'IAruptured == "no"')
round(prop.table(table(sim)), 2)

sim = cpdist(madbn_fit, nodes = c("gender", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
barplot(prop.table(table(sim)),
     main = 'IAruptured == "yes"')

sim = cpdist(madbn_fit, nodes = c("gender", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
barplot(prop.table(table(sim)),
     main = 'IAruptured == "no"')

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-location-gender"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```

```{r}
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAruptured"), n = 10^4, evidence = (smoking_current_notcurrent == "current" & hpt_aware == "yes" & positive_famillial_history == "yes"))
plot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "current", hpt_aware = "yes", positive_famillial_history = "yes"')
sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAruptured"), n = 10^4, evidence = (smoking_current_notcurrent == "not_current" & hpt_aware == "no" & positive_famillial_history == "no"))
plot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "not_current" & hpt_aware == "no" & positive_famillial_history == "no"')

sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAruptured"), n = 10^4, evidence = (smoking_current_notcurrent == "current" & hpt_aware == "yes" & positive_famillial_history == "yes"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "current", hpt_aware = "yes", positive_famillial_history = "yes"')
sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAruptured"), n = 10^4, evidence = (smoking_current_notcurrent == "not_current" & hpt_aware == "no" & positive_famillial_history == "no"))
barplot(prop.table(table(sim)), col = "grey",
     main = 'smoking_current_notcurrent == "not_current" & hpt_aware == "no" & positive_famillial_history == "no"')

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-age-smoking-hbp-family"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH*1.2, height = PLOTWIDTH)
    dev.off()
  } 
}
```

```{r Corresponding to Morel et al. 2022, Fig 3a-b }
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("multipleIAs", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(multipleIAs, IAlocation_group | IAruptured == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("multipleIAs", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(multipleIAs, IAlocation_group | IAruptured == 'no')")
prop.table(table(sim))

# if (SAVEPLOTS){
#   PLOTNAME <- "infquerry_rupture-multiple-location_barplt"
#   if(PLOTFORMAT == "svg"){
#     dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } else if(PLOTFORMAT == "png"){
#     dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } else if(PLOTFORMAT == "pdf"){
#     dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } 
# }
# 
# par(mfrow=c(1,2))
sim = cpdist(madbn_fit, nodes = c("multipleIAs", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
plot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(multipleIAs, IAlocation_group | IAruptured == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("multipleIAs", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
plot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(multipleIAs, IAlocation_group | IAruptured == 'no')")
prop.table(table(sim))

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-multiple-location"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```

```{r Corresponding to Morel et al. 2022, Fig 3c}
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(IAsize_diag_grouped_merged, IAlocation_group | IAruptured == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(IAsize_diag_grouped_merged, IAlocation_group | IAruptured == 'no')")
prop.table(table(sim))

# if (SAVEPLOTS){
#   PLOTNAME <- "infquerry_rupture-multiple-location_barplt"
#   if(PLOTFORMAT == "svg"){
#     dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } else if(PLOTFORMAT == "png"){
#     dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } else if(PLOTFORMAT == "pdf"){
#     dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } 
# }
# 
# par(mfrow=c(1,2))
sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
plot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(IAsize_diag_grouped_merged, IAlocation_group | IAruptured == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("IAsize_diag_grouped_merged", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
plot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(IAsize_diag_grouped_merged, IAlocation_group | IAruptured == 'no')")
prop.table(table(sim))

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-size-location"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```

```{r Corresponding to Morel et al. 2022, Fig 3d}
par(mfrow=c(2,2))
sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(age_diag_group, IAlocation_group | IAruptured == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(age_diag_group, IAlocation_group | IAruptured == 'no')")
prop.table(table(sim))

# if (SAVEPLOTS){
#   PLOTNAME <- "infquerry_rupture-multiple-location_barplt"
#   if(PLOTFORMAT == "svg"){
#     dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } else if(PLOTFORMAT == "png"){
#     dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } else if(PLOTFORMAT == "pdf"){
#     dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
#     dev.off()
#   } 
# }
# 
# par(mfrow=c(1,2))
sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "yes"))
plot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(age_diag_group, IAlocation_group | IAruptured == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("age_diag_group", "IAlocation_group"), n = 10^4, evidence = (IAruptured == "no"))
plot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(age_diag_group, IAlocation_group | IAruptured == 'no')")
prop.table(table(sim))

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_rupture-age-location"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTWIDTH)
    dev.off()
  } 
}
```

```{r}
par(mfrow=c(1,2))
sim = cpdist(madbn_fit, nodes = c("multipleIAs"), n = 10^4, evidence = (hpt_aware == "yes"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(multipleIAs | hpt_aware == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("multipleIAs"), n = 10^4, evidence = (hpt_aware == "no"))
barplot(prop.table(table(sim)), col = "grey",
     ylim = c(0,0.8), 
     main = "P(multipleIAs | hpt_aware == 'no')")
prop.table(table(sim))

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_multiple_hbp"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } 
}
```


```{r}
par(mfrow=c(1,3))
sim = cpdist(madbn_fit, nodes = c("multipleIAs"), n = 10^4, evidence = (IAlocation_group == "low"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.8),
     main = "P(multipleIAs | IAlocation_group == 'low')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("multipleIAs"), n = 10^4, evidence = (IAlocation_group == "medium"))
barplot(prop.table(table(sim)), col = "grey",
     ylim = c(0,0.8), 
     main = "P(multipleIAs | IAlocation_group == 'medium')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("multipleIAs"), n = 10^4, evidence = (IAlocation_group == "high"))
barplot(prop.table(table(sim)), col = "grey",
     ylim = c(0,0.8), 
     main = "P(multipleIAs | IAlocation_group == 'high')")
prop.table(table(sim))

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_multiple_location"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } 
} 
```


```{r}
par(mfrow=c(1,2))
sim = cpdist(madbn_fit, nodes = c("IAlocation_group"), n = 10^4, evidence = (hpt_aware == "yes"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.5), 
     main = "P(IA Location | hpt_aware == 'yes')")
prop.table(table(sim))

sim = cpdist(madbn_fit, nodes = c("IAlocation_group"), n = 10^4, evidence = (hpt_aware == "no"))
barplot(prop.table(table(sim)), col = "grey",
        ylim = c(0,0.5),
     main = "P(IA Location | hpt_aware == 'no')")
prop.table(table(sim))

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_location_hbp"
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } 
}
```
```{r }
par(mfrow = c(1,2))
sim = cpdist(madbn_fit, nodes = c("age_diag_group"), n = 10^4, evidence = (hpt_aware == "yes"))
plot(sim, col = "grey",
     main = "hpt_aware == yes")

sim = cpdist(madbn_fit, nodes = c("age_diag_group"), n = 10^4, evidence = (hpt_aware == "no"))
plot(sim, col = "grey",
     main = "hpt_aware == no")

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_hpt-age"  
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } 
}
```

```{r}
str(data)
```

# with restricted size->location

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(dplyr)
library(bnlearn)


# Save plots as files? 
SAVEPLOTS <- F
PLOTPATH <- paste0(Sys.getenv("PLOTPATH"), "/fourierplots")
PLOTFORMAT <- "pdf" # "pdf", "svg" or "png"
PLOTWIDTH = 16
PLOTHEIGHT = 9
```

## Load discrete additive mixed-effects BN

```{r message=FALSE, warning=FALSE}
load(paste0(PLOTPATH, "/discrete_BN_SL_additive_bic_mixedeffects_restrictedSizeLoc.RData"))
```


```{r}
data <- tobesaved$madbn_data %>%
  mutate(positive_famillial_history = factor(positive_famillial_history))
bl <- tobesaved$madbn_bl
avg <- tobesaved$madbn_avg
bb <- tobesaved$madbn_boot
```

```{r}
strength.plot(avg,
              bb,
              main = "discrete additive mixed-effects BN",
              # sub = "SL: tabu\nth.=0.3\nScore=BIC",
              shape = "rectangle")
```



## Inference by querying


```{r}
madbn_fit <- bn.fit(data = data, x = avg)
```

```{r}
# remove study_source
avg_mod <- bnlearn::remove.node(avg, "study_source")
bb_mod <-  bb %>%
  filter(from != "study_source") %>%
    filter(to != "study_source")
attr(bb_mod, "nodes") <- attr(bb_mod, "nodes")[-10]
data_mod <- data %>%
  select(-c("study_source"))

strength.plot(avg_mod,
              bb_mod,
              main = "discrete additive mixed-effects BN",
              sub = "Fitting w/o study_source",
              shape = "rectangle")

if (SAVEPLOTS){
  PLOTNAME <- "infquerry_admeBN_fitting_wo_studysource"  
  if(PLOTFORMAT == "svg"){
    dev.print(svg, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "png"){
    dev.print(png, filename = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } else if(PLOTFORMAT == "pdf"){
    dev.print(pdf, file = paste0(PLOTPATH, "/", PLOTNAME, ".", PLOTFORMAT), width = PLOTWIDTH, height = PLOTHEIGHT)
    dev.off()
  } 
} 

madbn_fit_mod <- bn.fit(data = data_mod, x = avg_mod)
```
