---
title: "Model Validation of discrete BN"
author: "Matteo Delucchi"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 3
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{Model Validation of discrete BN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r development, eval=FALSE, include=FALSE}
# devtools::document()
# devtools::load_all()
# renv::snapshot(prompt = F)
```

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(tidyr)
library(dplyr)
library(ggplot2)
library(parallel)
library(bnlearn)
library(forcats)

# Save plots as files? 
SAVEPLOTS <- F
PLOTPATH <- Sys.getenv("PLOTPATH")
PLOTFORMAT <- "png" # "svg" or "png"
```

# Load discrete BNs

The object `discrete_bns` was generated in vignette `discrete_BN_SL`.

```{r message=FALSE, warning=FALSE}
data <- exp11_dat$abndata
bl <- exp11_dat$bl
avgnet <- discrete_bns$avgnet
arcstren <- discrete_bns$arcstren
# str(avgnet)
```

```{r}
strength.plot(avgnet[["tabu.th03"]],
              arcstren[["tabu"]],
              main = "DAG",
              sub = "SL: tabu\nth.=0.3\nScore=BIC",
              shape = "rectangle")
strength.plot(avgnet[["tabu.bic.custom"]],
              arcstren[["tabu.bic.custom"]],
              main = "DAG",
              sub = "SL: tabu\nth.=sign.level\nScore=BIC-custom",
              shape = "rectangle")
```

# Cross-validation for predictive accuracy

## Classification Error of rupture status.

```{r}
xval.rupture.bic.part <- bn.cv(data = data, 
              bn = avgnet$tabu,
              method = "hold-out",
              runs = 50,
              fit = "mle",
              loss = "pred",
              loss.args = list(target="Ruptured_IA"))
xval.rupture.bic <- bn.cv(
    data = data,
    bn = avgnet$tabu.th03,
    method = "hold-out",
    runs = 50,
    fit = "mle",
    loss = "pred",
    loss.args = list(target = "Ruptured_IA")
  )
xval.rupture.bic.custom <- bn.cv(
    data = data,
    bn = avgnet$tabu.bic.custom,
    method = "hold-out",
    runs = 50,
    fit = "mle",
    loss = "pred",
    loss.args = list(target =  "Ruptured_IA")
  )
cv.metrics(xval.rupture.bic.part)
cv.metrics(xval.rupture.bic)
cv.metrics(xval.rupture.bic.custom)
```

```{r}
if (SAVEPLOTS) {
  FILE <- paste0(PLOTPATH, "/", "classification_error_ruptureStatus_all", ".")
  if (PLOTFORMAT == "svg") {
    svg(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "png") {
    png(paste0(FILE, PLOTFORMAT))
  }
  plot(xval.rupture.bic.part, xval.rupture.bic, xval.rupture.bic.custom,
     xlab = c("BIC.signTH", "BIC.th03", "BIC.custom"),
     main = "Classification Error of Rupture Status")
  dev.off()
} else {
  plot(xval.rupture.bic.part, xval.rupture.bic, xval.rupture.bic.custom,
     xlab = c("BIC.signTH", "BIC.th03", "BIC.custom"),
     main = "Classification Error of Rupture Status")
}
```


## Averaged, Overall Classification Error of Rupture Status

Average predictive accuracy over all nodes. 

```{r}
xval.list.bic <- list()
for (node in nodes(avgnet$tabu.th03)) {
  xval <- bn.cv(
    data = data,
    bn = avgnet$tabu.th03,
    method = "hold-out",
    runs = 50,
    fit = "mle",
    loss = "pred",
    loss.args = list(target = node)
  )
  xval.list.bic[[node]] <- list(xval = xval, 
                            cv.metrics = cv.metrics(xval))
}


xval.list.bic.custom <- list()
for (node in nodes(avgnet$tabu.th03)) {
  xval <- bn.cv(
    data = data,
    bn = avgnet$tabu.bic.custom,
    method = "hold-out",
    runs = 50,
    fit = "mle",
    loss = "pred",
    loss.args = list(target = node)
  )
  xval.list.bic.custom[[node]] <- list(xval = xval, 
                            cv.metrics = cv.metrics(xval))
}
accs.bic <- unlist(lapply(unlist(xval.list.bic, recursive = F), `[[`, "acc"))
summary(accs.bic, na.rm = T)
accs.bic.custom <- unlist(lapply(unlist(xval.list.bic.custom, recursive = F), `[[`, "acc"))
summary(accs.bic.custom, na.rm = T)
```


BIC.th03 used three parents (pos.fam.hist, size, location) to predict rupture
whereas BIC.custom uses only the later two parent nodes to predict rupture.
From `?bnlearn::bn.cv()`:  
"...  
Classification Error (pred): the prediction error for a single node in a discrete network. Frequentist predictions are used, so the values of the target node are predicted using only the information present in its local distribution (from its parents). Lower values are better.  

Posterior Classification Error (pred-lw and pred-lw-cg): similar to the above, but predictions are computed from an arbitrary set of nodes using likelihood weighting to obtain Bayesian posterior estimates. pred-lw applies to discrete Bayesian networks, pred-lw-cg to (discrete nodes in) hybrid networks. Lower values are better.  
..."  

## Posterior Classification Error of rupture status

So, we can use a Bayesian predictions that use a random set of available nodes for the prediction.

```{r}
xval.bic.bayes.part <- bn.cv(data = data, 
              bn = avgnet$tabu,
              method = "hold-out",
              runs = 50,
              fit = "mle",
              loss = "pred-lw",
              loss.args = list(target="Ruptured_IA"))
xval.bic.bayes <- bn.cv(data = data, 
              bn = avgnet$tabu.th03,
              method = "hold-out",
              runs = 50,
              fit = "mle",
              loss = "pred-lw",
              loss.args = list(target="Ruptured_IA"))
xval.bic.custom.bayes <- bn.cv(data = data, 
              bn = avgnet$tabu.bic.custom,
              method = "hold-out",
              runs = 50,
              fit = "mle",
              loss = "pred-lw",
              loss.args = list(target="Ruptured_IA"))
cv.metrics(xval.bic.bayes.part)
cv.metrics(xval.bic.bayes)
cv.metrics(xval.bic.custom.bayes)
```

```{r}
if (SAVEPLOTS) {
  FILE <- paste0(PLOTPATH, "/", "post_classification_error_ruptureStatus_all", ".")
  if (PLOTFORMAT == "svg") {
    svg(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "png") {
    png(paste0(FILE, PLOTFORMAT))
  }
  plot(xval.bic.bayes.part, xval.bic.bayes, xval.bic.custom.bayes, 
     xlab = c("BIC.signTH", "BIC.th03", "BIC.custom"),
     main = "Posterior Classification Error of Rupture Status")
  dev.off()
} else {
  plot(xval.bic.bayes.part, xval.bic.bayes, xval.bic.custom.bayes, 
     xlab = c("BIC.signTH", "BIC.th03", "BIC.custom"),
     main = "Posterior Classification Error of Rupture Status")
}
```

# Inference by querying

```{r}
tabu_th03_fit <- bn.fit(data = data, x = avgnet$tabu.th03)
```


```{r}
plot(tabu_th03_fit$location.grouped$prob)
```

Males tend to have an IA more often at a high risk location but
Females tend to develop an IA in high risk location more often if they suffer HBP than males who suffer HBP.


```{r}
plot(tabu_th03_fit$IAsize.groups.merged$prob)
```

Being aware of pos. fam. hist. reduces the porbability of developing larger IAs.

```{r}
(tabu_th03_fit$Ruptured_IA$prob)
```

Larger IAs tend to rupture more frequently.

```{r cpdist_age_hypertension_rupture}
par(mfrow = c(1,2))
sim = cpdist(tabu_th03_fit, nodes = c("Hypertension", "AgeDiag.group"), n = 10^4, evidence = (Ruptured_IA == "Yes"))
plot(sim, col = "grey",
     main = "Ruptured_IA == Yes")

sim = cpdist(tabu_th03_fit, nodes = c("Hypertension", "AgeDiag.group"), n = 10^4, evidence = (Ruptured_IA == "No"))
plot(sim, col = "grey",
     main = "Ruptured_IA == No")

par(mfrow = c(1,2))
sim = cpdist(tabu_th03_fit, nodes = c("AgeDiag.group", "Hypertension"), n = 10^4, evidence = (Ruptured_IA == "Yes"))
plot(sim, col = "grey",
     main = "Ruptured_IA == Yes")

sim = cpdist(tabu_th03_fit, nodes = c("AgeDiag.group", "Hypertension"), n = 10^4, evidence = (Ruptured_IA == "No"))
plot(sim, col = "grey",
     main = "Ruptured_IA == No")

par(mfrow = c(1,2))
sim = cpdist(tabu_th03_fit, nodes = c("AgeDiag.group"), n = 10^4, evidence = (Hypertension == "AnyType"))
plot(sim, col = "grey",
     main = "Hypertension == AnyType")

sim = cpdist(tabu_th03_fit, nodes = c("AgeDiag.group"), n = 10^4, evidence = (Hypertension == "Never"))
plot(sim, col = "grey",
     main = "Hypertension == Never")

par(mfrow = c(1,2))
sim = cpdist(tabu_th03_fit, nodes = c("AgeDiag.group"), n = 10^4, evidence = (Ruptured_IA == "Yes"))
plot(sim, col = "grey",
     main = "Ruptured_IA == Yes")

sim = cpdist(tabu_th03_fit, nodes = c("AgeDiag.group"), n = 10^4, evidence = (Ruptured_IA == "No"))
plot(sim, col = "grey",
     main = "Ruptured_IA == No")
```



```{r cpquery_smoking_gender}
plot(tabu_th03_fit$Smoking_Current_Former_No$prob)
```

Less former smokers than current or non-smokers.
Males more frequently are current smokers and females more frequently are former smokers.
Males most often stopped smoking.
-> Due to prior IA diagnosis?
-> Due to IA rupture?
-> Due to pos. fam. history?

```{r}
sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No"), n = 10^4, evidence = (Ruptured_IA == "Yes"))
plot(sim, col = "grey",
     main = 'Ruptured_IA == "Yes"')
sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No"), n = 10^4, evidence = (Ruptured_IA == "No"))
plot(sim, col = "grey",
     main = 'Ruptured_IA == "No"')
```

P(smoking | ruptured = Y)
P(smoking | ruptured = N)

```{r}
sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No"), n = 10^4, evidence = (Multiple.IAs == "Yes" & Gender == "Male"))
plot(sim, col = "grey",
     main = 'Multiple.IAs == "Yes" & Gender == "Male"')

sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No", "Multiple.IAs"), n = 10^4, evidence = (Gender == "Male"))
plot(sim, col = "grey",
     main = 'Gender == "Male"')

sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No", "Ruptured_IA"), n = 10^4, evidence = (Gender == "Male"))
plot(sim, col = "grey",
     main = 'Gender == "Male"')

sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No", "Positive.famillial.history"), n = 10^4, evidence = (Gender == "Male"))
plot(sim, col = "grey",
     main = 'Gender == "Male"')

sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No"), n = 10^4, evidence = (Multiple.IAs == "Yes" & Gender == "Male" & Ruptured_IA == "Yes"))
plot(sim, col = "grey",
     main = 'Multiple.IAs == "Yes" & Gender == "Male" & Ruptured_IA == "Yes"')

sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No"), n = 10^4, evidence = (Gender == "Male" & Ruptured_IA == "Yes"))
plot(sim, col = "grey",
     main = 'Multiple.IAs == "Yes" & Gender == "Male" & Ruptured_IA == "Yes"')
```


```{r}
par(mfrow=c(1,2))
sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No", "Ruptured_IA"), n = 10^5, evidence = (Gender == "Female"))
plot(sim, col = "grey",
     main = 'Gender == "Female"')

sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No", "Ruptured_IA"), n = 10^5, evidence = (Gender == "Male"))
plot(sim, col = "grey",
     main = 'Gender == "Male"')
```

Former Smokers have a slightly increased rupture risk compared to the others.

```{r}
sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No", "Ruptured_IA"), n = 10^4, evidence = (Positive.famillial.history == "Yes"))
plot(sim, col = "grey")
table(sim)

# sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No"), n = 10^4, evidence = (Positive.famillial.history == "No" & Ruptured_IA == "Yes"))
# plot(sim, col = "grey")
# prop.table(table(sim))
# sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No"), n = 10^4, evidence = (Positive.famillial.history == "Yes" & Ruptured_IA == "Yes"))
# plot(sim, col = "grey")
# prop.table(table(sim))
```

But being aware of IA cases in relatives decreases rupture risk of former smokers.


```{r}
sim = cpdist(tabu_th03_fit, nodes = c("Smoking_Current_Former_No", "Hypertension", "Positive.famillial.history"), n = 10^4, evidence = (Ruptured_IA == "Yes"))
plot(sim, col = "grey")
prop.table(table(sim))
```

```{r}
sim = cpdist(tabu_th03_fit, nodes = c("Ruptured_IA"), n = 10^4, evidence = (Multiple.IAs == "Yes" ))
prop.table(table(sim))
plot(sim)

sim = cpdist(tabu_th03_fit, nodes = c("Ruptured_IA"), n = 10^4, evidence = (Multiple.IAs == "No" ))
prop.table(table(sim))
plot(sim)
```

No clear tendency for rupture with and withouth multiple IAs.

```{r}
sim = cpdist(tabu_th03_fit, nodes = c("Ruptured_IA", "IAsize.groups.merged"), n = 10^4, evidence = (Multiple.IAs == "Yes" ))
prop.table(table(sim))
plot(sim)

sim = cpdist(tabu_th03_fit, nodes = c("Multiple.IAs", "IAsize.groups.merged"), n = 10^4, evidence = (Ruptured_IA == "Yes" ))
prop.table(table(sim))
plot(sim)
```

Having multiple IAs, they tend to rupture at larger size.


```{r}
sim = cpdist(tabu_th03_fit, nodes = c("Ruptured_IA"), n = 10^4, evidence = (Gender == "Female" & location.grouped == "High"))
prop.table(table(sim))

sim = cpdist(tabu_th03_fit, nodes = c("Ruptured_IA", "IAsize.groups.merged"), n = 10^4, evidence = (Gender == "Female" & Hypertension == "AnyType" & Multiple.IAs == "Yes" ))
prop.table(table(sim))
```

```{r}
str(data)
```

