---
title: "Structure Learning of discrete BN"
author: "Matteo Delucchi"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 3
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{Structure Learning of discrete BN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r development, eval=FALSE, include=FALSE}
devtools::document()
devtools::load_all()
# renv::snapshot(prompt = F)
```

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(tidyr)
library(dplyr)
library(ggplot2)
library(parallel)
library(bnlearn)
library(forcats)

# Save plots as files? 
# Warning: Bad layout with knitr. Works well if chunks are run without rendering.
SAVEPLOTS <- FALSE
PLOTPATH <- Sys.getenv("PLOTPATH")
```

# Load Preprocessed Data

```{r message=FALSE, warning=FALSE}
data <- exp11_dat$abndata
bl <- exp11_dat$bl
str(data)
```


```{r}
arcstren <- list()
avgnet <- list()
avgnet.th0 <- list()
```

# Bootstrap size optimization

We estimate the size of the non-parametric bootstrap replicate stepwise to 
investigate on the the model performance and number of bootstrap replicates.

Since it is an open question to compare an estimated model to the true unknown model
(see [(Thesis of Scutari 2011)](https://www.bnlearn.com/about/thesis/thesis.pdf)), 
we try different measures.

I don't go into detail about all the applied measures at this point. Please 
look in the help page of the functions in `structureanalysis.R` for more details.

```{r bootstrap eval, echo=FALSE, message=FALSE, warning=FALSE}
cl = makeCluster(6)

eval <- NULL

for (b in c(100,seq(500, 10000, 1000))) {
  print("--------------------------------------------------------------------------------")
  tabusize = 18
  crt = "bic" # BIC standard
  priorname = "uniform" # uniform standard
  algo = "tabu"
  print(paste("Bootstrap replicates: ", b))

  boot.tabu <- boot.strength(data = data,
                      R = b,
                        algorithm = algo,
                        algorithm.args = list(
                          tabu = tabusize,
                          blacklist = bl,
                          score = crt,
                          prior = priorname),     # prior distribution to be used with the various Bayesian Dirichlet scores
                        cluster = cl)

  avg.boot.tabu <- averaged.network(boot.tabu)
  avg.boot.tabu.th0 <- averaged.network(boot.tabu, threshold = 0)


  netmet <- network.metrics(data, avg.boot.tabu, avg.boot.tabu.th0, algo=algo, tabulistsize=tabusize, b=b, trueGraph=ekg, crt=crt, priorname=priorname)

  if (!is.null(eval)){
    eval <- rbind(eval, netmet)
  } else{
    eval <- netmet
    }
  }
stopCluster(cl)

eval.bootstrap.rep.tabu <- eval

eval.bootstrap.rep.tabu.plot.data <- eval.bootstrap.rep.tabu %>%
  prep.data2plot() %>%
  select(-c(tabu.list.size, Algorithm)) %>%
  data.table::melt(id.vars = c("bootstrap.replicates", "score.function", "prior")) %>%
  mutate(across(5, function(x){round(as.numeric(x),3)})) %>%

  # rearrange order (of bootstrap.replicates) for grouped display (by prior) in graph
  mutate(bootstrap.replicates = as.integer(bootstrap.replicates)) %>%
  arrange(desc(bootstrap.replicates), .group_by = TRUE) %>%

  ggplot() +
  aes(x = bootstrap.replicates, y = value, group = score.function, color = prior)+
  geom_point() +
  # geom_text(aes(label=round(value,2)), hjust=0, vjust=2, angle = -45)+
  geom_line()+
  facet_grid(c("variable"), scales = "free_y")+
  # facet_wrap(~variable, scales = "free_y") +
  # scale_x_continuous(breaks = c(100,500,1000,5000))+
  theme_minimal()+
  scale_color_discrete(name = "Prior",
                       labels = c("Uniform"))+
#   scale_y_continuous(breaks = function(x, n = 5) {
#     # Show only integers on y-axis
#     l <- pretty(x, n)
#     ifelse(!all(l > 0),
#       return(l[abs(l %% 1) < .Machine$double.eps ^ 0.5]),
#       return(l))
#     })+
  ggtitle(c("Structure Learning Metrics"),
        subtitle = paste("Varying Bootstrap Replicas.\nSL Algorithm: ", algo,"\nTabulist size: ", tabusize, "\nScore:", crt)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text.y = element_text(angle = 0, hjust = 0),
        panel.background = element_rect(),
        panel.spacing = unit(1, "lines"))
```


```{r classic bootstrap plot, warning=FALSE, layout="l-screen", out.height=2400, out.width="80%"}
eval.bootstrap.rep.tabu.plot.data +
  theme(panel.spacing = unit(0.3, "lines"))
```


The network score is the typical goodness of fit measure of such a model. 
We do not detect a clear relationship. Therefore, and because of still reasonable
computation time, we just assume a large number of bootstrap replicates.

# Structure learning algorithm assessment

Analogous to the non-parametric bootstrap size estimate above, it would be 
interesting to measure the model performance resulting from a set of different
structure learning algorithms.  

The open issue of measuring the goodness of fit of a model stays the same as 
when optimizing the number bootstrap replicas: it's very difficult to measure and 
in practice the network score (mostly BIC) is used to compare model performance.

The results are therefore omitted at this point.

# Structure Learning with TABU search and BIC

For aforementioned reasons, we focus on recent but standard methodologies 
and parameters.

The final DAG is the averaged structural model resulting from repetitively applying
the structure learning algorithm TABU on a non-parametric bootstrap sample. 

```{r bootstrap data, echo=FALSE, message=FALSE, warning=FALSE}
cl = makeCluster(6)

# Structure learning with non-parametric bootstrap framework
cl = makeCluster(6)
algo = "tabu"
b = 10000
tabusize = 18
crt = "bic" # BIC standard
priorname = "uniform" # uniform standard
boot <- boot.strength(
  data = data,
  R = b,
  algorithm = algo,
  algorithm.args = list(
    tabu = tabusize,
    blacklist = bl,
    score = crt,
    prior = priorname
  ),
  # prior distribution to be used with the various Bayesian Dirichlet scores
  cluster = cl
)
stopCluster(cl)

arcstren[[algo]] <- boot

avg.boot <- averaged.network(boot)
avgnet[[algo]] <- avg.boot

avg.boot.th0 <- averaged.network(boot, threshold = 0)
avgnet.th0[[algo]] <- avg.boot.th0

netmet <- network.metrics(
  data,
  estGraph = avg.boot,
  estGraph.th0 = avg.boot.th0,
  algo = algo,
  tabulistsize = ifelse(algo == "tabu", tabusize, NA),
  b = b,
  trueGraph = ekg,
  crt = crt,
  priorname = priorname
)
```


```{r}
plot(arcstren[["tabu"]])
abline(v=0.3)
```


We see many arcs of low support and no visually clear cut-point to set the arc-strength
threshold. 

We see can see it is difficult to argue from a statistical point of view
where to set the threshold and this is a reason why we have the significance 
threshold as a point of orientation.

The significance threshold is close to the 50% threshold and the structure would
not change upon increasing to that point.

If it is reasonable one can vary from the proposed significance threshold.


```{r cpdag-tabu-strength}
strength.plot(cpdag(avg.boot, wlbl = TRUE),
              arcstren[["tabu"]],
              main = "CPDAG",
              sub = "SL: tabu\nth.=sign.level",
              shape = "rectangle")
```

As already expected from the arc-strength plot, we see a very sparse, partially
connected CPDAG with no arc directions.

Variables, we know they must have an effect on rupture or at least some other
variables are not connected.

Eventhough methodologically sound, it is difficult to justify this network from
a clinical point of view. 

Interestingly all connections from the CPDAG above were found in the [consensus DAG
above](#mcmcconsensusDAG) as well. 



If we lower the arc-strength threshold to 30% (dashed line in arc-strength plot. 
We stay in the allowed range of variation) we result in a fully connected, 
partially directed CPDAG.

```{r cpdag-tabu-strength-th30, layout="l-page"}
avgnet[["tabu.th03"]] <- averaged.network(arcstren[["tabu"]], threshold = 0.3)

strength.plot(cpdag(avgnet[["tabu.th03"]], wlbl = TRUE),
              arcstren[["tabu"]],
              main = "CPDAG",
              sub = "SL: tabu\nth.=0.3",
              shape = "rectangle")
```


```{r}
# svg(paste0(PLOTPATH, "/dag_tabu_th03.svg"))
strength.plot(avgnet[["tabu.th03"]],
              arcstren[["tabu"]],
              main = "DAG",
              sub = "SL: tabu\nth.=0.3",
              shape = "rectangle")
# dev.off()
# bnlearn::write.dot(graph = dag.th03, file = paste0(PLOTPATH, "/dag_tabu_th03.dot"))
# write.csv(arcstren[["tabu"]], file = paste0(PLOTPATH, "/arc_strengths.csv"))
```

# Structure Learning with TABU search and custom BIC


```{r bootstrap bic custom, echo=FALSE, message=FALSE, warning=FALSE}
cl = makeCluster(6)

# Structure learning with non-parametric bootstrap framework
cl = makeCluster(6)
algo = "tabu"
b = 10000
tabusize = 18
crt = "bic" # BIC standard
priorname = "uniform" # uniform standard
boot <- boot.strength(
  data = data,
  R = b,
  algorithm = algo,
  algorithm.args = list(
    tabu = tabusize,
    blacklist = bl,
    score = crt,
    prior = priorname,
    k = log(nrow(data)) / 4
  ),
  # prior distribution to be used with the various Bayesian Dirichlet scores
  cluster = cl
)
stopCluster(cl)

arcstren[["tabu.bic.custom"]] <- boot

avg.boot <- averaged.network(boot)
avgnet[["tabu.bic.custom"]] <- avg.boot

avg.boot.th0 <- averaged.network(boot, threshold = 0)
avgnet.th0[["tabu.bic.custom"]] <- avg.boot.th0

netmet <- network.metrics(
  data,
  estGraph = avg.boot,
  estGraph.th0 = avg.boot.th0,
  algo = algo,
  tabulistsize = ifelse(algo == "tabu", tabusize, NA),
  b = b,
  trueGraph = ekg,
  crt = crt,
  priorname = priorname
)
```


```{r}
plot(arcstren[["tabu.bic.custom"]])
abline(v=0.3)
```



```{r cpdag-tabu-strength-bic-custom}
strength.plot(cpdag(avg.boot, wlbl = TRUE),
              arcstren[["tabu.bic.custom"]],
              main = "CPDAG",
              sub = "SL: tabu\nth.=sign.level\nScore=BIC-custom",
              shape = "rectangle")
```


```{r}
# svg(paste0(PLOTPATH, "/dag_tabu_th03.svg"))
strength.plot(avgnet[["tabu.bic.custom"]],
              arcstren[["tabu.bic.custom"]],
              main = "DAG",
              sub = "SL: tabu\nth.=sign.level\nScore=BIC-custom",
              shape = "rectangle")
# dev.off()
# bnlearn::write.dot(graph = dag.th03, file = paste0(PLOTPATH, "/dag_tabu_th03.dot"))
# write.csv(arcstren[["tabu"]], file = paste0(PLOTPATH, "/arc_strengths.csv"))
```



## save to package data

```{r save avgnet and arcstren}
discrete_bns <- list(avgnet = avgnet, 
                     arcstren = arcstren)
usethis::use_data(discrete_bns, overwrite = TRUE)
devtools::document()
```
