---
title: "Addressing Reviewers Comments"
author: "Matteo Delucchi"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 3
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{Structure Learning of discrete BN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r development, eval=FALSE, include=FALSE}
# devtools::document()
# devtools::load_all()
# renv::snapshot(prompt = F)
```

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(tidyr)
library(dplyr)
library(ggplot2)
library(parallel)
library(bnlearn)
library(forcats)

# Save plots as files? 
SAVEPLOTS <- F
PLOTPATH <- Sys.getenv("PLOTPATH")
PLOTFORMAT <- "pdf" # "svg", "pdf" or "png"
```


# Prepare data set


```{r}
str(adb)
```

## Individual Location groups

Age: discrete  
Location: As individual parent vessel
Size: discrete and merged (see vignette `data_preprocessing` for details)  

```{r}
expBV_disc_dat <- prep_exp_data(
  dat = adb,
  EXPNO = 88,
  age = "disc.grouped-multinomial",
  location = "byVessel-multinomial",
  size = "grouped.merged-multinomial",
  smoking = "mult",
  SAVE = F)
```

The prepared data is returned as list of inputs for the following 
BN modeling:

```{r}
disc_data <- expBV_disc_dat$abndata
disc_dist <- expBV_disc_dat$dist
disc_banned <- expBV_disc_dat$banned
disc_bl <- expBV_disc_dat$bl

str(disc_data)
str(disc_dist)
str(disc_banned)
str(disc_bl)
```

```{r}
expBV_cont_dat <- prep_exp_data(
  dat = adb,
  EXPNO = 99,
  age = "cont",
  location = "byVessel-multinomial",
  size = "log",
  smoking = "mult",
  SAVE = F)
```

The prepared data is returned as list of inputs for the following 
BN modeling:

```{r}
cont_data <- expBV_cont_dat$abndata
cont_dist <- expBV_cont_dat$dist
cont_banned <- expBV_cont_dat$banned
cont_bl <- expBV_cont_dat$bl

str(cont_data)
str(cont_dist)
str(cont_banned)
str(cont_bl)
```

# DBN SL

```{r}
arcstren <- list()
avgnet <- list()
avgnet.th0 <- list()
```

## Structure Learning with TABU search and BIC

```{r bootstrap data, echo=FALSE, message=FALSE, warning=FALSE}
cl = makeCluster(6)

# Structure learning with non-parametric bootstrap framework
cl = makeCluster(6)
algo = "tabu"
b = 10000
tabusize = 18
crt = "bic" # BIC standard
priorname = "uniform" # uniform standard
boot <- boot.strength(
  data = disc_data,
  R = b,
  algorithm = algo,
  algorithm.args = list(
    tabu = tabusize,
    blacklist = disc_bl,
    score = crt,
    prior = priorname
  ),
  # prior distribution to be used with the various Bayesian Dirichlet scores
  cluster = cl
)
stopCluster(cl)

arcstren[[algo]] <- boot

avg.boot <- averaged.network(boot)
avgnet[[algo]] <- avg.boot

avg.boot.th0 <- averaged.network(boot, threshold = 0)
avgnet.th0[[algo]] <- avg.boot.th0

netmet <- network.metrics(
  disc_data,
  estGraph = avg.boot,
  estGraph.th0 = avg.boot.th0,
  algo = algo,
  tabulistsize = ifelse(algo == "tabu", tabusize, NA),
  b = b,
  trueGraph = NULL,
  crt = crt,
  priorname = priorname
)
```



```{r}
if (SAVEPLOTS) {
  FILE <- paste0(PLOTPATH, "/", "arcstren_tabu_bic_vesselloc", ".")
  if (PLOTFORMAT == "svg") {
    svg(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "png") {
    png(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "pdf") {
    pdf(paste0(FILE, PLOTFORMAT))
  }
  plot(arcstren[["tabu"]])
  # abline(v=0.3)
  dev.off()
} else {
  plot(arcstren[["tabu"]])
  abline(v = 0.3)
}
```




```{r cpdag-tabu-strength}
if (SAVEPLOTS) {
  FILE <- paste0(PLOTPATH, "/", "cpdag_tabu_bic_vesselloc", ".")
  if (PLOTFORMAT == "svg") {
    svg(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "png") {
    png(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "pdf") {
    pdf(paste0(FILE, PLOTFORMAT))
  }
  strength.plot(
    cpdag(avg.boot, wlbl = TRUE),
    arcstren[["tabu"]],
    main = "CPDAG",
    sub = "SL: tabu\nth.=sign.level",
    shape = "rectangle"
  )
  dev.off()
} else {
  strength.plot(
    cpdag(avg.boot, wlbl = TRUE),
    arcstren[["tabu"]],
    main = "CPDAG",
    sub = "SL: tabu\nth.=sign.level",
    shape = "rectangle"
  )
}
```

```{r dag-tabu-strength}
if (SAVEPLOTS) {
  FILE <- paste0(PLOTPATH, "/", "dag_tabu_bic_vesselloc", ".")
  if (PLOTFORMAT == "svg") {
    svg(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "png") {
    png(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "pdf") {
    pdf(paste0(FILE, PLOTFORMAT))
  }
  strength.plot(
    avgnet[["tabu"]],
    arcstren[["tabu"]],
    main = "DAG",
    sub = "SL: tabu\nth.=sign.level",
    shape = "rectangle"
  )
  dev.off()
} else {
  strength.plot(
    avgnet[["tabu"]],
    arcstren[["tabu"]],
    main = "DAG",
    sub = "SL: tabu\nth.=sign.level",
    shape = "rectangle"
  )
}
```

```{r cpdag-tabu-strength-th30, layout="l-page"}
avgnet[["tabu.th03"]] <- averaged.network(arcstren[["tabu"]], threshold = 0.3)


if (SAVEPLOTS) {
  FILE <- paste0(PLOTPATH, "/", "cpdag_tabu_bic_th03_vesselloc", ".")
  if (PLOTFORMAT == "svg") {
    svg(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "png") {
    png(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "pdf") {
    pdf(paste0(FILE, PLOTFORMAT))
  }
  strength.plot(
    cpdag(avgnet[["tabu.th03"]], wlbl = TRUE),
    arcstren[["tabu"]],
    main = "CPDAG",
    sub = "SL: tabu\nth.=0.3",
    shape = "rectangle"
  )
  dev.off()
} else {
  strength.plot(
    cpdag(avgnet[["tabu.th03"]], wlbl = TRUE),
    arcstren[["tabu"]],
    main = "CPDAG",
    sub = "SL: tabu\nth.=0.3",
    shape = "rectangle"
  )
}
```

The respective DAG:

```{r}
if (SAVEPLOTS) {
  FILE <- paste0(PLOTPATH, "/", "dag_tabu_bic_th03_vesselloc", ".")
  if (PLOTFORMAT == "svg") {
    svg(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "png") {
    png(paste0(FILE, PLOTFORMAT))
  } else if (PLOTFORMAT == "pdf") {
    pdf(paste0(FILE, PLOTFORMAT))
  }
  strength.plot(avgnet[["tabu.th03"]],
                arcstren[["tabu"]],
                main = "DAG",
                sub = "SL: tabu\nth.=0.3",
                shape = "rectangle")
  dev.off()
  bnlearn::write.dot(graph = avgnet[["tabu.th03"]],
                     file = paste0(FILE, "dot"))
  write.csv(arcstren[["tabu"]], file = paste0(FILE, "csv"))
} else {
  strength.plot(avgnet[["tabu.th03"]],
                arcstren[["tabu"]],
                main = "DAG",
                sub = "SL: tabu\nth.=0.3",
                shape = "rectangle")
  }
```
