---
title: "Harmonizing Raw Data"
author: "Matteo Delucchi"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 3
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{Harmonizing Raw Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Clear working environment
rm(list=ls())

# Load libraries
library(bnaiaR)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(RMariaDB)
library(forcats)

# Save plots as files? 
# Warning: Bad layout with knitr. Works well if chunks are run without rendering.
SAVEPLOTS <- FALSE
PLOTPATH <- Sys.getenv("PLOTPATH")
```

# Raw Data

Raw data is stored in two SQL data bases which are not public. 

## Dependencies

This file connects to the database docker container published on port 6603 on
`localhost`.
The instructions to set up the container stack are available on:
[`github.zhaw.ch/delt/aneurysmDB`](https://github.zhaw.ch/delt/aneurysmDB) in a 
private, unpublished repo.

```{r debug, include=FALSE}
# Secrets are stored in unpublished environment variables.
USERNAME <- Sys.getenv("USERNAME")
PASSWORD <- Sys.getenv("PASSWORD")
DB_CONFIG <- Sys.getenv("DB_CONFIG")
```


We set up the connection to the DBs.

```{r warning=FALSE}
# Connect to my-db as defined in ~/.my.cnf
conaneux_CRIM_ANEUQUEST_2021_06 <- connectDBstack(dbname="aneux_CRIM_ANEUQUEST_2021_06",
                      username=USERNAME,
                      password=PASSWORD,
                      default.file = DB_CONFIG)
conaneux_CRIM_ANEUX_2021_06 <- connectDBstack(dbname="aneux_CRIM_ANEUX_2021_06",
                      username=USERNAME,
                      password=PASSWORD,
                      default.file = DB_CONFIG)

# All connection variable names in one vector
cons <- ls(pattern = "^con")
```

Query the raw data from the two data bases: aneuquest and aneux.

```{r}
aneuquest.raw <- dbGetQuery(conaneux_CRIM_ANEUQUEST_2021_06, "
SELECT
    ROOT.patientID,
    ROOT.patDOB,
    ROOT.aneuristID,
    ROOT.recruBasis,
    ROOT.patSex,
    ROOT.patEthnicity,
    ROOT.prf_PosFamHis,
    ROOT.prf_SmokedTobacco,
    ROOT.prf_HyperTension,
    aneu.ID,
    aneu.ruptureStatus,
    aneu.aneuLoca,
    aneu.maxDiam,
    aneu.aneuReportDate,
    aneu.aneuReportPatAge,
    aneu.aneuUniqueID    
FROM
    ROOT
LEFT JOIN EventID AS ev
ON
    ROOT.ID = ev.ParentID
RIGHT JOIN ANEUDETAILEVENT AS aneu
ON
      ev.ID = aneu.ParentID
")


aneux.raw <- dbGetQuery(conaneux_CRIM_ANEUX_2021_06, "
SELECT
    ROOT.patientID,
    ROOT.patDOB,
    ROOT.aneuristID,
    ROOT.recruBasis,
    ROOT.patSex,
    ROOT.patEthnicity,
    ROOT.prf_PosFamHis,
    ROOT.prf_SmokedTobacco,
    ROOT.prf_HyperTension,
    aneu.ID,
    aneu.ruptureStatus,
    aneu.aneuLoca,
    aneu.maxDiam,
    aneu.aneuReportDate,
    aneu.initialDiagDate,
    aneu.aneuReportPatAge,
    aneu.noAnevrisme,
    aneu.aneuUniqueID
FROM
    ROOT
LEFT JOIN EventID AS ev
ON
    ROOT.ID = ev.ParentID
RIGHT JOIN ANEUDETAILEVENT AS aneu
ON
    ev.ID = aneu.ParentID
")
```

```{r}
str(aneuquest.raw)
str(aneux.raw)
```

## Harmonization

```{r}
aneuquest <- crim4toISGC(aneuquest.raw) 
```

```{r}
aneux <- crim5toISGC(aneux.raw) 
```

## Transformation 
### Add reference to data base source
For later reference we add a column with the name of the data source
```{r}
aneuquest[["db"]] <- "aneuquest"
aneux[["db"]] <- "aneux"
```

### Join all ADB together

```{r}
adb.raw <- aneuquest %>%
  bind_rows(aneux) 
```

Take care about the patient anonymisation by removing the patient hash.

```{r}
identifier <- paste0("pat", sample(1:1000000, nrow(adb.raw))) 

adb.raw <- adb.raw %>%
  mutate(ID_2 = identifier) 
```


The raw data then looks like this:

```{r}
str(adb.raw)
```

The number of entries by data base:

```{r}
table(adb.raw$db)
```

## Validation

```{r, fig.height=14, out.height= "100%"}
plt <- overview_plt_max(adb.raw)

if (SAVEPLOTS){
  ggsave(plot = plt, filename = "ADB_overview.png", path = PLOTPATH, dpi = 600)}
```

### Age

```{r}
table(adb.raw$age.at.time.of.diagnosis, useNA = "always")
```

Filter impossible age values:

```{r}
patientsWithWrongAge <- adb.raw %>%
  filter(age.at.time.of.diagnosis > 130 | is.na(age.at.time.of.diagnosis) | age.at.time.of.diagnosis <= 0) %>%
  select(ID_2, age.at.time.of.diagnosis)
```


### Size

```{r}
table(adb.raw$IA_size.at.time.of.SAH, useNA = "always")
```


```{r}
summary(as.numeric(adb.raw$IA_size.at.time.of.SAH))
```

## save ADB raw data set to R package

### Make all variables factors
```{r}
adb.raw <- adb.raw %>%
  # convert all variables to factors
  mutate(across(where(purrr::is_character), forcats::as_factor),
         # across(where(is_numeric),as_factor)
         ) %>%
  # order
  mutate(Positive.famillial.history = forcats::fct_relevel(Positive.famillial.history, c("Yes","No")))%>%
  mutate(Smoking_Current_Former_No = forcats::fct_relevel(Smoking_Current_Former_No, c("Current","Former","No")))%>%
  mutate(Ruptured_IA = forcats::fct_relevel(Ruptured_IA, c("Yes","No")))%>%
  mutate(Multiple.IAs = forcats::fct_relevel(Multiple.IAs, c("Yes", "No")))%>%
  mutate(IA_size.at.time.of.SAH.groups = forcats::fct_relevel(IA_size.at.time.of.SAH.groups, LETTERS[1:length(levels(IA_size.at.time.of.SAH.groups))]))%>%
  mutate(Age.at.diag.grouped = forcats::fct_relevel(Age.at.diag.grouped, LETTERS[1:length(levels(Age.at.diag.grouped))])) %>%
  mutate(Hypertension = forcats::fct_relevel(Hypertension, c("AnyType", "Never")))%>%
  mutate(IA_Location = forcats::fct_relevel(IA_Location, c("Acom", "Pcom","V-B", "A2","PC","MCA","ICA","Basilar","Other","A1 segment ant","CavICA","OphtICA")))%>%
  mutate(Gender = forcats::fct_relevel(Gender, c("Female", "Male"))) %>%
  # as data frame
  as.data.frame()
```

## Save ADB_raw as RData

```{r}
str(as.data.frame(adb.raw))
```

To make this data set available in other vignettes, we store it in `inst/extdata`.

```{r save adb_raw}
# save(adb.raw, file = "../inst/extdata/adb_raw.RData")
```

The data set can then be loaded via:

```{r}
# rm(list=ls())
# load(system.file("extdata", "adb_raw.RData", package = "bnaiaR"))
```

## Include Preprocessed Data in bnaiaR package

```{r}
# usethis::use_data(adb.raw, overwrite = TRUE)
# devtools::document()
```

