---
title: "Regression Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regression Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  out.height = "100%",
  fig.width = 12,
  fig.height = 9
)
```

```{r development, eval=FALSE, include=FALSE}
# devtools::document()
# devtools::load_all()
# renv::snapshot(prompt = F)

```


```{r setup}
# Clear working environment
rm(list=ls())

library(bnaiaR)
library(dplyr)
library(ggplot2)
library(forcats)
library(mgcv)

set.seed(100)

SAVEPLOTS <- F
PLOTPATH <- Sys.getenv("PLOTPATH")
PLOTWIDTH = 16
PLOTHEIGHT = 9
```

## Load Data

```{r load_data}
data(exp11_dat)
data = exp11_dat$abndata
str(data)
```

# Multiple Logistic Regression Model 

```{r}
data.reg <- data %>%
  mutate(location.grouped = fct_relevel(location.grouped, c("Low","Medium","High")))%>%
  mutate(Smoking_Current_Former_No = fct_relevel(Smoking_Current_Former_No, c("No","Former","Current"))) %>%
  mutate(Multiple.IAs = fct_relevel(Multiple.IAs, c("No", "Yes"))) %>%
  mutate(Gender = fct_relevel(Gender,c("Male","Female"))) %>%
  mutate(Hypertension = fct_relevel(Hypertension,c("Never","AnyType"))) %>%
  mutate(Ruptured_IA = fct_relevel(Ruptured_IA,c("No","Yes"))) %>%
  mutate(Positive.famillial.history = fct_relevel(Positive.famillial.history, c("No", "Yes"))) %>%
  mutate(AgeDiag.group = fct_relevel(AgeDiag.group, c(LETTERS[seq(8,1,-1)]))) %>%
  mutate(IAsize.groups.merged = fct_relevel(IAsize.groups.merged, c(LETTERS[seq(3,1,-1)])))
```



## Log reg for IA Rupture

As in Figure 6 in Morel et al. 2022.

```{r}
data.logreg <- data.reg %>%
  select(c(location.grouped, Smoking_Current_Former_No, Multiple.IAs, Gender, Hypertension, Ruptured_IA)) 
```


### Data preparation

```{r}
train <- sample.int(n = nrow(data.logreg), size = 0.7*nrow(data.logreg), replace = FALSE)
TrainSet <- data.logreg[train,]
ValidSet <- data.logreg[-train,]
str(TrainSet)
summary(TrainSet)
str(ValidSet)
summary(ValidSet)
```



```{r}
glm_model <- glm(Ruptured_IA ~.,family=binomial(link='logit'),data=TrainSet)
summary (glm_model)

plt <- sjPlot::plot_model(glm_model,
           show.values = TRUE, 
           value.offset = .3, 
           sort.est=TRUE,  
           line.size = 1.5, 
           dot.size = 3, #
           # transform="plogis", # plots probabilities
           # value.size	= 4,
           title = "Logistic regression for IA rupture in ADB") +
  ylab("Odds ratios") +
  geom_hline(yintercept = 1) +
  theme_light()

if (SAVEPLOTS){
  plt.mod <- plt +
    sjPlot::font_size(
                      title = 15,
                      axis_title.y = 15,
                      axis_title.x = 15,
                      labels.x = 15,
                      labels.y = 15,
                      base.theme = theme_light())
  ggsave(plot = plt.mod, filename = "logreg_IArupture.png", path = PLOTPATH, 
         # dpi = 600, width = 4096, height = 3072, units = "px")
         dpi = 600, width = 4096, height = 2304, units = "px")
} else {
  plt
}
```


```{r}
glm_model
ANOVA <- anova(glm_model,test="Chisq")
ANOVA$`Pr(>Chi)`
ANOVA


pscl::pR2(glm_model)
pred_glm <- predict(glm_model,newdata=ValidSet,type='response')
table(ValidSet$Ruptured_IA, pred_glm > 0.5)

roc <- pROC::roc(ValidSet$Ruptured_IA, pred_glm)
auc <- roc$auc
aucCI <- round(pROC::ci.auc(ValidSet$Ruptured_IA, pred_glm), 2)
plotROCAUC(roc, aucCI, FILENAME = "logreg_IArupture_roc.png", PLOTPATH = PLOTPATH)

pred_glm.rocr <- ROCR::prediction(pred_glm, ValidSet$Ruptured_IA)
perf.acc <- ROCR::performance(pred_glm.rocr, measure="acc", x.measure = "cutoff")
ROCR::plot(perf.acc, main = "Auccuracy")
# abline(v = 0.5)

perf.precrec <- ROCR::performance(pred_glm.rocr, measure="prec", x.measure="rec")
plot(perf.precrec, 
     main = "Precision/recall")

paste("Accuracy = ", max(perf.acc@y.values[[1]]))
perf.acc@x.values[[1]][which(perf.acc@y.values[[1]] == max(perf.acc@y.values[[1]]))]
# perf.acc@y.values[[1]][max(which(perf.acc@x.values[[1]] >= 0.5))]

## Accuraccy as in Morel et al 2021 for verification:
# why 0.48?
pred_glm <- ifelse(pred_glm > 0.53,"Yes","No")
misClasificError <- mean(pred_glm != ValidSet$Ruptured_IA)
format(misClasificError, digits = 2)
lp_accuracy_glm <- (paste("Accuracy = ",1-misClasificError))
lp_accuracy_glm
```


## Log reg for IA Rupture including all Variables

```{r}
data.logreg.all <- data.reg %>%
  rename(c("Sex" = "Gender",
           # "Sex = Female" = "Gender.Female",
           # "Sex = Male" = "Gender.Male",
           "Fam. History" = "Positive.famillial.history",
           # "Fam. History = No" = "Positive.famillial.history.No",
           # "Fam. History = Yes" = "Positive.famillial.history.Yes",
           "Age" = "AgeDiag.group",           
           # "Age = F" = "AgeDiag.group.F",
           # "Age = G" = "AgeDiag.group.G",
           # "Age = D" = "AgeDiag.group.D",
           # "Age = E" = "AgeDiag.group.E",
           # "Age = C" = "AgeDiag.group.C",
           # "Age = B" = "AgeDiag.group.B",
           # "Age = A" = "AgeDiag.group.A",
           "Hypertension" = "Hypertension",
           # "Hypertension = Never" = "Hypertension.Never",
           # "Hypertension = AnyType" = "Hypertension.AnyType",
           "Smoking" = "Smoking_Current_Former_No",
           # "Smoking = Current" = "Smoking_Current_Former_No.Current",
           # "Smoking = Former" = "Smoking_Current_Former_No.Former",
           # "Smoking = No" = "Smoking_Current_Former_No.No",
           "Location" = "location.grouped",
           # "Location = High risk" = "location.grouped.High",
           # "Location = Medium risk" = "location.grouped.Medium",
           # "Location = Low risk" = "location.grouped.Low",
           "IA size" = "IAsize.groups.merged",
           # "IA size = A" = "IAsize.groups.merged.A",
           # "IA size = B" = "IAsize.groups.merged.B",
           # "IA size = C" = "IAsize.groups.merged.C",
           "Multiple IAs" = "Multiple.IAs",
           # "Multiple IAs = Yes" = "Multiple.IAs.Yes",
           # "Multiple IAs = No" = "Multiple.IAs.No",
           "Ruptured IA" = "Ruptured_IA"
           # "Ruptured IA = Yes" = "Ruptured_IA.Yes",
           # "Ruptured IA = No" = "Ruptured_IA.No"
           ))
```

### Data preparation

```{r}
train <- sample.int(n = nrow(data.logreg.all), size = 0.7*nrow(data.logreg.all), replace = FALSE)
TrainSet <- data.logreg.all[train,]
ValidSet <- data.logreg.all[-train,]
```

Cross-validation

```{r message=FALSE, warning=FALSE}
# defining training control
# as Leave One Out Cross Validation
train_control <- caret::trainControl(method = "LOOCV",
                                     summaryFunction = caret::prSummary, 
                                     # summaryFunction = twoClassSummary, 
                                     classProbs = TRUE,
                                     savePredictions = "all")
 
# training the model by assigning sales column
# as target variable and rest other column
# as independent variable
model <- caret::train(`Ruptured IA` ~., data = data.logreg.all,
               method = "glm",
               family=binomial(link='logit'),
               trControl = train_control)
 
# printing model performance metrics
# along with other details
print(model)

loocv.cm.logreg <- caret::confusionMatrix(model$pred$pred, model$pred$obs, positive = "Yes", mode = "everything")
df.predmetrics <- data.frame(loocv.logreg = round(c(loocv.cm.logreg$overall, loocv.cm.logreg$byClass), 3))
df.predmetrics
```

70:30 split

```{r}
glm_model.logreg.all <- glm(`Ruptured IA` ~.,family=binomial(link='logit'),data=TrainSet)
summary (glm_model.logreg.all)

plt <- sjPlot::plot_model(glm_model.logreg.all,
           show.values = TRUE, 
           value.offset = .3, 
           sort.est=TRUE,  
           line.size = 1.5, 
           dot.size = 3, #,transform="plogis" plots probabilities
           title = "Logistic regression for IA rupture") +
  ylab("Odds ratios") +
  geom_hline(yintercept = 1) +
  theme_light() 

if (SAVEPLOTS){
  plt.mod <- plt +
    sjPlot::font_size(
                      title = 15,
                      axis_title.y = 15,
                      axis_title.x = 15,
                      labels.x = 15,
                      labels.y = 15,
                      base.theme = theme_light())
  ggsave(plot = plt.mod, filename = "logreg_IArupture_allvars.png", path = PLOTPATH, 
         dpi = 600, width = 4096, height = 4096, units = "px")
         # dpi = 600, width = 4096, height = 2304, units = "px")
} else {
  plt
}

sjPlot::tab_model(glm_model.logreg.all)
```



```{r}
glm_model.logreg.all
ANOVA.logreg.all <- anova(glm_model.logreg.all,test="Chisq")
ANOVA.logreg.all$`Pr(>Chi)`
ANOVA.logreg.all


pscl::pR2(glm_model.logreg.all)
pred_glm.logreg.all <- predict(glm_model.logreg.all,newdata=ValidSet,type='response')
table(ValidSet$`Ruptured IA`, pred_glm.logreg.all > 0.5)


conf.mat <- table(as.numeric(ValidSet$`Ruptured IA`), as.numeric(pred_glm.logreg.all>0.5))
colnames(conf.mat) <- rownames(conf.mat) <- c("No", "Yes")

split.cm.logreg.all <- caret::confusionMatrix(conf.mat, mode = "everything", positive = "Yes")
df.predmetrics <- cbind(df.predmetrics, data.frame(split.logreg = round(c(split.cm.logreg.all$overall, split.cm.logreg.all$byClass), 3)))
df.predmetrics

roc.logreg.all <- pROC::roc(ValidSet$`Ruptured IA`, pred_glm.logreg.all)
aucCI.logreg.all <- round(pROC::ci.auc(ValidSet$`Ruptured IA`, pred_glm.logreg.all), 2)
plotROCAUC(roc.logreg.all, aucCI.logreg.all, FILENAME = "logreg_IArupture_allvars_roc.png", PLOTPATH = PLOTPATH)

pred.logreg.all.rocr <- ROCR::prediction(pred_glm.logreg.all, ValidSet$`Ruptured IA`)
perf.acc.logreg.all <- ROCR::performance(pred.logreg.all.rocr, measure="acc", x.measure = "cutoff")
ROCR::plot(perf.acc.logreg.all, main = "Auccuracy")

perf.precrec.logreg.all <- ROCR::performance(pred.logreg.all.rocr, measure="prec", x.measure="rec")
plot(perf.precrec.logreg.all, 
     main = "Precision/recall")

paste("Accuracy = ", max(perf.acc.logreg.all@y.values[[1]]))
max.acc.cutoff <- perf.acc.logreg.all@x.values[[1]][which(perf.acc.logreg.all@y.values[[1]] == max(perf.acc.logreg.all@y.values[[1]]))]
paste("Cutoff with max. Accuracy = ", max.acc.cutoff)
# perf.acc.logreg.all@y.values[[1]][max(which(perf.acc.logreg.all@x.values[[1]] >= 0.5))]

## Accuraccy as in Morel et al 2021 for verification:
pred_glm.logreg.all <- ifelse(pred_glm.logreg.all > max.acc.cutoff,"Yes","No")
misClasificError.logreg.all <- mean(pred_glm.logreg.all != ValidSet$`Ruptured IA`)
# format(misClasificError.logreg.all, digits = 2)
lp_accuracy_glm.logreg.all <- (paste("Accuracy = ",1-misClasificError.logreg.all))
lp_accuracy_glm.logreg.all
```

## Log reg for IA Rupture including all Variables but age

```{r}
data.logreg.all.noage <- data.reg %>%
  select(-c(AgeDiag.group))
```

### Data preparation

```{r}
train <- sample.int(n = nrow(data.logreg.all.noage), size = 0.7*nrow(data.logreg.all.noage), replace = FALSE)
TrainSet <- data.logreg.all.noage[train,]
ValidSet <- data.logreg.all.noage[-train,]
```



```{r}
glm_model.logreg.all.noage <- glm(Ruptured_IA ~.,family=binomial(link='logit'),data=TrainSet)
summary(data.logreg.all.noage)

plt <- sjPlot::plot_model(glm_model.logreg.all.noage,
           show.values = TRUE, 
           value.offset = .3, 
           sort.est=TRUE,  
           line.size = 1.5, 
           dot.size = 3, #,transform="plogis" plots probabilities
           title = "Logistic regression for IA rupture in ADB (all variables)") +
  ylab("Odds ratios") +
  geom_hline(yintercept = 1) +
  theme_light() 

if (SAVEPLOTS){
  plt.mod <- plt +
    sjPlot::font_size(
                      title = 15,
                      axis_title.y = 15,
                      axis_title.x = 15,
                      labels.x = 15,
                      labels.y = 15,
                      base.theme = theme_light())
  ggsave(plot = plt, filename = "logreg_IArupture_allvars.png", path = PLOTPATH, 
         dpi = 600, width = 4096, height = 4096, units = "px")
         # dpi = 600, width = 4096, height = 2304, units = "px")
} else {
  plt
}
```



# Multivariate Generalized Additive Model for IA Rupture including all variables

Replacing categorical `AgeDiag.group` with continuous `AgeDiag`.

```{r}
gam_dat <- adb %>%
  select(c(Gender, 
           AgeDiag, 
           # AgeDiag.group, 
           Positive.famillial.history, Smoking_Current_Former_No, Hypertension, location.grouped, Multiple.IAs,
           # IAsize.groups, 
           IAsize_log, 
           Ruptured_IA)) %>%
  # Remove IA size outliers
  filter(IAsize_log < 4) %>%
  # Clean up age values
  filter(AgeDiag != 0) %>%
  mutate(location.grouped = fct_relevel(location.grouped, c("Low","Medium","High")))%>%
  mutate(Smoking_Current_Former_No = fct_relevel(Smoking_Current_Former_No, c("No","Former","Current"))) %>%
  mutate(Multiple.IAs = fct_relevel(Multiple.IAs, c("No", "Yes"))) %>%
  mutate(Gender = fct_relevel(Gender,c("Male","Female"))) %>%
  mutate(Hypertension = fct_relevel(Hypertension,c("Never","AnyType"))) %>%
  mutate(Ruptured_IA = fct_relevel(Ruptured_IA,c("No","Yes"))) %>%
  mutate(Positive.famillial.history = fct_relevel(Positive.famillial.history, c("No", "Yes"))) 
  # mutate(AgeDiag.group = fct_relevel(AgeDiag.group, c(LETTERS[seq(8,1,-1)]))) %>%
  # mutate(IAsize.groups.merged = fct_relevel(IAsize.groups.merged, c(LETTERS[seq(3,1,-1)])))
  # mutate(Ruptured_IA = as.integer(Ruptured_IA))

str(gam_dat)
```

### Data preparation

```{r}
train <- sample.int(n = nrow(gam_dat), size = 0.7*nrow(gam_dat), replace = FALSE)
TrainSet <- gam_dat[train,]
ValidSet <- gam_dat[-train,]
```

### Generalized additive model

```{r}
# gam_model <- gam(Ruptured_IA ~ Positive.famillial.history+location.grouped+s(IAsize_log, k=3)+s(AgeDiag, k=3), data=TrainSet, family = "binomial")
gam_model <- gam(Ruptured_IA ~ Gender+Positive.famillial.history+Hypertension+Smoking_Current_Former_No+location.grouped+s(IAsize_log, k=3)+Multiple.IAs+s(AgeDiag, k=3), data=TrainSet, family = "binomial")
gam.check(gam_model)
summary.gam(gam_model)
par(mfrow = c(1,8))
plot.gam(gam_model, residuals=T, se=TRUE, 
         shade=T, shade.col='gray90', 
         all.terms = T, pages = 1)
abline(h=0)

anova.gam(gam_model)
```

```{r}
if (SAVEPLOTS){
  graphics.off()

  PLOTNAME <- "gam_model"
  pdf(file = paste0(PLOTPATH, "/", PLOTNAME, ".pdf"), width = PLOTWIDTH*0.7, height = PLOTHEIGHT)
  
  op <- par(mfrow=c(3,3))
            # , mai=c(0.01,0.01,0.01,0.01), mar=c(0.1,0.1,0.1,0.1))
  for (i in 1:9) {
    plot.gam(gam_model, select = i, residuals=T, se=TRUE,
             shade=T, shade.col='gray90',
             all.terms = T)
    abline(h=0)
    }
  par(op)
  
  dev.off()
} else {
  op <- par(mfrow=c(3,3))
  for (i in 1:9) {
    plot.gam(gam_model, select = i, residuals=T, se=TRUE, 
             shade=T, shade.col='gray90', 
             all.terms = T)  
    abline(h=0)
    }
  par(op)
}
```

```{r}
gam_pred <- predict.gam(gam_model, newdata = ValidSet, type = "response")
conf.mat <- table(ValidSet$Ruptured_IA, as.numeric(gam_pred>0.5))
colnames(conf.mat) <- rownames(conf.mat) <- c("No", "Yes")

split.cm.gam <- caret::confusionMatrix(conf.mat, mode = "everything", positive = "Yes")
df.predmetrics <- cbind(df.predmetrics, data.frame(split.gam = round(c(split.cm.gam$overall, split.cm.gam$byClass), 3)))
df.predmetrics

roc.gam_model <- pROC::roc(ValidSet$Ruptured_IA, gam_pred)
aucCI.gam_model <- round(pROC::ci.auc(ValidSet$Ruptured_IA, gam_pred), 2)
plotROCAUC(roc.gam_model, aucCI.gam_model, FILENAME = "gam_IArupture_allvars_roc.png", PLOTPATH = PLOTPATH) # TODO: CHeck this function!
```

Cross-validation gam

```{r message=FALSE, warning=FALSE}
# defining training control
# as Leave One Out Cross Validation
train_control <- caret::trainControl(method = "LOOCV",
                                     summaryFunction = caret::prSummary, 
                                     # summaryFunction = twoClassSummary, 
                                     classProbs = TRUE,
                                     savePredictions = "all")
 
# training the model by assigning sales column
# as target variable and rest other column
# as independent variable
model <- caret::train(Ruptured_IA ~ Gender+Positive.famillial.history+Hypertension+Smoking_Current_Former_No+location.grouped+IAsize_log+Multiple.IAs+AgeDiag, data = gam_dat,
               method = "gam",
               family=binomial(link='logit'),
               trControl = train_control,
               tuneGrid = data.frame(method = "GCV.Cp", select = FALSE))
 
# printing model performance metrics
# along with other details
print(model)

loocv.cm.gam <- caret::confusionMatrix(model$pred$pred, model$pred$obs, positive = "Yes", mode = "everything")
df.predmetrics <- cbind(df.predmetrics, data.frame(loocv.gam = round(c(loocv.cm.gam$overall, loocv.cm.gam$byClass), 3)))
df.predmetrics
```

```{r}
gam_model_summary <- summary(gam_model)
pltdat <- data.frame(varnames = names(gam_model_summary$p.coeff), coefs = as.numeric(gam_model_summary$p.coeff), pvalue=as.numeric(gam_model_summary$p.pv)) %>%
  mutate(expcoefs = exp(coefs))

ggplot(pltdat)+
  aes(x=varnames, y=coefs)+
  geom_point()+
  ylab("Logit") +
  geom_hline(yintercept = 0) +
  theme_light() +
  coord_flip()

ggplot(pltdat)+
  aes(x=varnames, y=expcoefs)+
  geom_point()+
  ylab("Odds ratios") +
  geom_hline(yintercept = 1) +
  theme_light() +
  coord_flip()

if (SAVEPLOTS){
  # plt.mod <- plt +
  #   sjPlot::font_size(
  #                     title = 15,
  #                     axis_title.y = 15,
  #                     axis_title.x = 15,
  #                     labels.x = 15,
  #                     labels.y = 15,
  #                     base.theme = theme_light())
  # ggsave(plot = plt, filename = "gam_IArupture_allvars.png", path = PLOTPATH, 
  #        dpi = 600, width = 4096, height = 4096, units = "px")
  #        # dpi = 600, width = 4096, height = 2304, units = "px")
} else {
  plt
}
```

store prediction metrics

```{r}
write.csv(df.predmetrics, file = paste0(PLOTPATH, "/prediction_metrics_regression.csv"))
```


